---
title: "Covid 19 Report"
author: "Rupa"
date: "2023-07-03"
output:
  pdf_document: default
  html_document: default
  word_document: default
papersize: a4
---

## Covid-19 Data

```{r setup, include = FALSE}

library (tidyverse)
library (lubridate)

``` 

## Introduction

Covid-19 is an infectious disease caused by the SARS virus. This virus causes acute respiratory syndrome. Most of the people globally experienced Covid-19, ranging from severe to mild. This report is related to the Covid-19 Data collected by Johns Hopkins. This report helps understand the Covid-19 pattern Globally and in the US.

## Import and Read Data

```{r}

# Get the Covid Data:

url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

file_names <- c("time_series_covid19_confirmed_global.csv", "time_series_covid19_deaths_global.csv",
"time_series_covid19_confirmed_US.csv",
"time_series_covid19_deaths_US.csv")

# Read the Covid Data:

urls <- str_c (url_in, file_names)
urls

global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
us_cases <- read_csv(urls[3])
us_deaths <- read_csv(urls[4])

```

## Tidy and Transform the data

# Clean the global cases and deaths data keeping only columns needed for the analysis and join the cases and deaths data

```{r}

# Clean the global cases and deaths data keeping only columns needed for the analysis

global_cases <- global_cases %>% pivot_longer (cols = -c('Province/State', 'Country/Region', Lat, Long), names_to = "date", values_to = "cases") %>% select (-c(Lat, Long))

global_deaths <- global_deaths %>% pivot_longer (cols = -c('Province/State', 'Country/Region', Lat, Long), names_to = "date", values_to = "deaths")%>% select (-c(Lat, Long))

# Join the global cases and deaths data

global <- global_cases %>% full_join (global_deaths) %>% rename (Country_Region = 'Country/Region', Province_State = 'Province/State')

# Extract the month and year from the date

global$date <- format(as.Date(global$date, format="%m/%d/%y"),"%m/%y")

# Filter out cases greater that 28000000

global %>% filter (cases > 28000000)

# Combine the Province/State and Country/Region to one column using a comma separator

global <- global %>% unite ("Combined_Key", c(Province_State, Country_Region), sep = " , ", na.rm = TRUE, remove = FALSE)

# Get the updated list of the global cases and deaths after cleaning up the data
# and keeping only columns needed for the analysis

global

```

# Clean the US cases and deaths data keeping only columns needed for the analysis and join the cases and deaths data

```{r}

us_cases <- us_cases %>% pivot_longer (cols = -(UID:Combined_Key), names_to = "date", values_to = "cases") %>% select(Admin2:cases) %>% select (-c(Lat, Long_))

us_deaths <- us_deaths %>% pivot_longer (cols = -(UID:Population), names_to = "date", values_to = "deaths") %>% select(Admin2:deaths) %>% mutate (date = mdy(date)) %>% select (-c(Lat, Long_))

# Extract the month and year from the date for the cases and deaths

us_cases$date <- format(as.Date(us_cases$date, format="%m/%d/%y"),"%m/%y")
us_deaths$date <- format(as.Date(us_deaths$date, format="%m/%d/%y"),"%m/%y")

# Join the US cases and deaths data

US <- us_cases %>% full_join (us_deaths)

# Get the updated list of the US cases and deaths after cleaning up the data
# and keeping only columns needed for the analysis

US

```

## Import the population data for all the countries globally and Read Data

```{r}

# Import the Data with the global population:

uid_lookup_url <-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

# Read the global population data:

uid <- read_csv (uid_lookup_url) %>% select (-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))

uid

```

## Tidy and Transform the data

# Clean the global population data keeping only columns needed for the analysis

```{r}

# Add the population data to the global data

global <- global %>% left_join (uid, by = c("Province_State",
"Country_Region")) %>% select (-c(UID, FIPS)) %>% select (Province_State, Country_Region, date, cases, deaths, Population, Combined_Key)

```

## Group the Global Data by Country/Region

```{r}

# Group the Global Data by Country/Region

global_by_country <- global %>% group_by (Country_Region, date) %>% summarize (cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>% mutate (deaths_per_mill = deaths * 1000000 / Population)

global_by_country <- global_by_country %>% select (Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% ungroup ()

global_by_country

```

## Plot the Global Cases Data

```{r}

# Plot the Global Data by cases

global_by_country %>% filter (cases > 0) %>% ggplot(aes (x = date, y = cases)) + geom_bar (stat = "identity", width = 0.7) + theme (legend.position = "bottom", axis.text.x = element_text (angle = 90)) + labs (title = "COVID19 cases globally", y = NULL)+ xlab ("Year") + ylab ("Cases")

```

## Plot the Global death Data

```{r}

# Plot the Global Data by deaths

global_by_country %>% filter (cases > 0) %>% ggplot(aes (x = date, y = deaths)) + geom_bar (stat = "identity", width = 0.7) + theme (legend.position = "bottom", axis.text.x = element_text (angle = 90)) + labs (title = "COVID19 deaths globally", y = NULL) + xlab ("Date") + ylab ("Deaths")

```

## Visualize the US data

```{r}

# Visualize the US Data by State

US_by_state <- US %>% group_by (Province_State, Country_Region, date) %>% summarize (cases = sum(cases), deaths = sum(deaths), Population = sum (Population)) %>% mutate (deaths_per_mill = deaths * 1000000 / Population)

US_by_state <- US_by_state %>% select (Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% ungroup ()

```

## Group the US Data by State by Country/Region

```{r}

# Group the US Data by State by Country/Region

US_totals <- US_by_state %>% group_by (Country_Region, date) %>% summarize (cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>% mutate (deaths_per_mill = deaths * 1000000 / Population)

US_totals <- US_by_state %>% select (Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% ungroup ()

US_totals
tail(US_totals)

```

## Plot the US Data

```{r}

# Plot the US Data

US_totals %>% filter (cases > 0) %>% ggplot(aes (x = date, y = cases)) + geom_line (aes (color = "cases")) + geom_point (aes (color = "cases")) + geom_line (aes (y = deaths, color = "deaths")) + geom_point (aes(y = deaths, color = "deaths")) + scale_y_log10() + theme (legend.position = "bottom", axis.text.x = element_text (angle = 90)) + labs (title = "COVID19 in US", y = NULL)

```

## Plot the Data for the State of New York

```{r}

# Plot the Data for the State of New York

state <- "New York"

US_by_state %>% filter (Province_State == state) %>% filter (cases >0) %>% ggplot (aes (x = date, y = cases)) + geom_line (aes (color = "cases")) + geom_point (aes (color = "cases")) + geom_line (aes (y= deaths, color = "deaths")) + geom_point (aes (y = deaths, color =
"deaths")) + scale_y_log10() + theme(legend.position = "bottom", axis.text.x = element_text (angle = 90)) + labs (title = str_c ("COVID19 in ", state), y = NULL)

```

## Analyze the data

```{r}

# Find the maximum date

max_date = max (US_totals$date)

max_date

# Find the maximum deaths

max_deaths = max (US_totals$deaths)

max_deaths

```

## Modify the data to add the new cases and new deaths

```{r}

# Modify the data to add the new cases and new deaths

US_by_state <- US_by_state %>% mutate (new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))

US_totals <- US_totals %>% mutate (new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))

US_totals

```

## Analyze the US_totals table

```{r}

# Check the tail end of the US_totals table

tail (US_totals)

# Check the new cases and new deaths from the tail end of the US_totals table

tail (US_totals %>% select (new_cases, new_deaths, everything()))

```

## Plot the US totals data

```{r}

# Plot the US totals data

US_totals %>% ggplot (aes (x = date, y = new_cases)) + geom_line (aes (color = "new_cases")) + geom_point (aes (color = "new_cases")) + geom_line (aes (y= new_deaths, color = "new_deaths")) + geom_point (aes (y = new_deaths, color = "new_deaths")) + scale_y_log10() + theme(legend.position = "bottom", axis.text.x = element_text (angle =
90)) + labs (title = str_c ("COVID19 in ", state), y = NULL)

```

## Plot the Data for the State of Colorado

```{r}

# Plot the Data for the State of Colorado

state <- "Colorado"

US_by_state %>% filter (Province_State == state) %>% filter (cases >0) %>% ggplot (aes (x = date, y = new_cases)) + geom_line (aes (color = "new_cases")) + geom_point (aes (color = "new_cases")) + geom_line (aes (y= new_deaths, color = "new_deaths")) + geom_point (aes (y = new_deaths, color = "new_deaths")) + scale_y_log10() + theme(legend.position = "bottom", axis.text.x = element_text (angle = 90)) + labs (title = str_c ("COVID19 in", state), y = NULL)

```

## Visualize the Data for all US States

```{r}

# Visualize the Data for all US States

US_state_totals <- US_by_state %>% group_by (Province_State) %>% summarize (deaths = max (deaths), cases = max (cases), population = max (Population), cases_per_thou = 1000 * cases / population, deaths_per_thou = 1000 * deaths / population) %>% filter (cases > 0,
population > 0)

# Get the list of 10 states with the minimum deaths per thousand

US_state_totals %>% slice_min (deaths_per_thou, n = 10) %>% select (deaths_per_thou, cases_per_thou, everything())

# Get the list of 10 states with the maximum deaths per thousand

US_state_totals %>% slice_max (deaths_per_thou, n = 10) %>% select (deaths_per_thou, cases_per_thou, everything())

US_state_totals

```

## Model the Data

```{r}

# Model the Data

mod <- lm (deaths_per_thou ~ cases_per_thou, data = US_state_totals)

summary (mod)

US_state_totals %>% slice_min (cases_per_thou)

US_state_totals %>% slice_max (cases_per_thou)

x_grid <- seq (100, 500)

new_df <- tibble (cases_per_thou = x_grid)

US_tot_w_pred <- US_state_totals %>% mutate (pred = predict(mod))

US_tot_w_pred

```

## Plot the Data

```{r}

# Plot the Data

US_tot_w_pred %>% ggplot () + geom_point (aes (x = cases_per_thou, y = deaths_per_thou), color = "blue") + geom_point (aes (x = cases_per_thou, y = pred), color = "red")

```

## Conclusion

The analysis shows how much Covid-19 affected everyone globally and the number of deaths caused by the Covid-19 virus.  It shows how much every country remained vulnerable to the virus.  Whether it is the vaccines or natural immunity that a person has towards Covid-19, it looks like there is finally a reduction in the number of cases and deaths related to the virus.