---
title: "NYPDShootingIncidentData"
author: "Rupa"
date: "2023-07-03"
output:
  pdf_document: default
  html_document: default
  word_document: default
papersize: a4
---

## NYPD Shooting Incident

```{r setup, include = FALSE}

library (tidyverse)
library (lubridate)

```

## Introduction

This report is related to the New York Shooting Incidents from 2006 to 2022. This report helps understand the pattern and trends in the shootings and the Boroughs that the shootings are most seen.

## Import and Read Data

```{r}

# Get the NYPD Shooting Incident Data:

ny_data <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv"

ny_cases <- read_csv (ny_data)

```

## Tidy and Transform the data

```{r}

# Clean the data keeping only columns needed for the analysis

ny_cases <- ny_cases %>% select (-c(JURISDICTION_CODE, OCCUR_TIME, LOC_OF_OCCUR_DESC,
                                    LOC_CLASSFCTN_DESC, LOCATION_DESC, PERP_AGE_GROUP,
                                    PERP_SEX, PERP_RACE, X_COORD_CD, Y_COORD_CD,
                                    Latitude, Longitude, Lon_Lat))

# Extract Unique Elements

ny_cases <- unique (ny_cases)

# Extract the year from the DATE

ny_cases$Year <- format(as.Date(ny_cases$OCCUR_DATE, format="%m/%d/%Y"),"%Y")

# Arrange data in Descending Order of Most Shootings

cases_by_boro <- ny_cases %>% group_by(BORO, Year) %>%  summarize (Cases = n())

# Get the updated list of the cases after cleaning up the data
# and keeping only columns needed for the analysis

cases_by_boro

```

## Plot the cases by Borough by Year

```{r}

# Plot the number of cases by Year by Borough

cases_by_boro %>% ggplot(aes (x = Year, y = Cases, fill = BORO)) + geom_bar (stat = "identity", width = 0.7) + xlab ("Year") + ylab ("Cases") + ggtitle ("New York Shooting Cases by Borough by Year") + theme (legend.position = "bottom") + scale_fill_manual(values=c('#d62828', '#8338ec', '#f77f00', '#0096c7', '#e75480'))

```

## Plot the cases by Borough

```{r}

# Plot the number of cases by Borough

cases_by_boro %>% ggplot(aes (x = BORO, y = Cases, fill = BORO)) + geom_bar (stat = "identity", width = 0.7) + xlab ("Borough") + ylab ("Cases") + ggtitle ("New York Shooting Cases by Borough") + theme (legend.position = "right") + scale_fill_manual(values=c('#d62828', '#8338ec', '#f77f00', '#0096c7', '#e75480'))

```

## Plot the cases by Victim Age Group

```{r}

# Plot the cases by Victim Age Group

cases_by_vic_age_group <- ny_cases %>% filter (VIC_AGE_GROUP != 1022)

cases_by_vic_age_group <- cases_by_vic_age_group %>% filter (VIC_AGE_GROUP != "UNKNOWN")

cases_by_vic_age_group_by_year <- cases_by_vic_age_group %>% group_by(VIC_AGE_GROUP, Year) %>%  summarize (Cases = n())

cases_by_vic_age_group_by_year[is.na(cases_by_vic_age_group_by_year) | cases_by_vic_age_group_by_year=="Inf"] = NA

cases_by_vic_age_group_by_year %>% ggplot(aes (x = VIC_AGE_GROUP, y = Cases, fill = VIC_AGE_GROUP)) + geom_bar (stat = "identity", width = 0.7) + xlab ("Victim Age Group") + ylab ("Cases") + ggtitle ("New York Shooting Cases by Victim Age Group") + theme (legend.position = "right") + scale_fill_manual(values=c('#d62828', '#8338ec', '#f77f00', '#0096c7', '#e75480'))

```

## Plot the cases by Victim Age Group by Year

```{r}

#Plot the cases by Victim Age Group by Year

cases_by_vic_age_group_by_year %>% ggplot(aes (x = Year, y = Cases, fill = VIC_AGE_GROUP)) + geom_bar (stat = "identity", width = 0.7) + xlab ("Year") + ylab ("Cases") + ggtitle ("New York Shooting Cases by Victim Age Group by Year") + theme (legend.position = "bottom") + scale_fill_manual(values=c('#d62828', '#8338ec', '#f77f00', '#0096c7', '#e75480'))

```

## Plot the cases by Victim Age Group by Borough

```{r}

#Plot the cases by Victim Age Group by Borough

cases_by_vic_by_boro <- cases_by_vic_age_group %>% group_by(Year, VIC_AGE_GROUP, BORO) %>% summarize (Cases = n())

cases_by_vic_by_boro

cases_by_vic_by_boro[is.na(cases_by_vic_by_boro) | cases_by_vic_by_boro=="Inf"] = NA

cases_by_vic_by_boro %>% ggplot(aes (x = VIC_AGE_GROUP, y = Cases, fill = BORO, label = Cases)) + geom_bar (stat = "identity", width = 0.7) + xlab ("Victim Age Group") + ylab ("Cases") + ggtitle ("New York Shooting Cases by Victim Age Group by Borough") + theme (legend.position = "right") + scale_fill_manual(values=c('#d62828', '#8338ec', '#f77f00', '#0096c7', '#e75480'))

```


## Linear Modeling the Cases by Victim's Age Group and Cases by Borough

```{r}

#Update the Cases by Borough to include Statistical Murder Analysis by Borough

total_cases <- ny_cases %>% summarize (boro_cases = cases_by_boro$Cases, vic_age_cases = cases_by_vic_age_group_by_year$Cases)

total_cases

lin_mod <-  lm (formula = boro_cases ~ vic_age_cases, data = total_cases)

summary (lin_mod)

total_cases %>% slice_min (vic_age_cases)
total_cases %>% slice_max (vic_age_cases)

total_cases

x_grid <- seq (-250, 250)
#lin_mod

lm_cases_with_pred <- total_cases %>% mutate (pred = predict (lin_mod))

lm_cases_with_pred %>% ggplot () + geom_point (aes (x = boro_cases, y = vic_age_cases), color = "blue") + geom_point (aes (x = boro_cases, y = pred), color = "red") + xlab ("Cases by Borough") + ylab ("Cases by Victim Age Group") + ggtitle ("Linear Modeling the Cases by Victim Age Group and Borough") + theme (legend.position = "bottom")

```

## Conclusion

The analysis shows that most of the cases are from Brooklyn and the lowest cases are from Staten Island. The rates had gone down between 2017-2019, but seems to have risen back in 2020. The reason could be because of the lower population of Staten Island compared to Brooklyn. However, not having strict gun laws and seeing the increase in shooting incidents is very concerning.
